---
title: "Blake Conrad -- CSCI 43200 Security -- Network Analysis -- Final Project"
output: html_notebook
---
<br><br>
<h2>Import the data</h2>
```{r}
library(ggplot2)
setwd("/Users/bmc/Desktop/CSCI-43200/finalProject/")
table <- read.csv("2013-10-01_capture-win12.conn_summary.csv", header=TRUE)
names(table)[8] <- "X.Security"
names(table)
```

<h2> One day's analysis as an example plot</h2>
```{r}
plt1 <- ggplot() + geom_line(data=table[1:6,c("Starttime","X.Packets","X.Security")], aes(x=Starttime, y=X.Packets,color=X.Security)) 
plt1 <- plt1 + ggtitle("Network Analysis", subtitle="Example of 1 day's security")
plt1 <- plt1 + xlab("Number of Hours (Cumulative by 5)") + ylab("Number of Packets")
plt1 <- plt1 + scale_colour_gradient(limits=c(1, 10), low="red", high="green")
plt1
```

<h2>Actually Network Security for the entire data set</h2>
<ul><li>Fluctuations and dead spots are getting the worst rankings for security</li></ul>
```{r}
plt2 <- ggplot() + geom_line(data=table[,c("Starttime","X.Packets","X.Security")], aes(x=Starttime, y=X.Packets,color=X.Security)) 
plt2 <- plt2 + ggtitle("Network Analysis", subtitle="Actual Network Security Results")
plt2 <- plt2 + xlab("Number of Hours (Cumulative by 5)") + ylab("Number of Packets")
plt2 <- plt2 + scale_colour_gradient(limits=c(0, 10), low="red", high="green")
plt2
```

<h2>Apply K-means on the data</h2>
<ul><li>At K=2</li></ul>
```{r}
set.seed(12345)
cluster <- kmeans(table[,2:ncol(table)], centers=2, nstart=30)
table$cluster <- cluster$cluster
head()
```

<h2>K-means Results</h2>
<ul>
<li>At K=2</li>
<li>The initial security vulnerabilies weren't captured</li>
<li>The dead spots and latter more severse secure vulnerabilities were captured</li>
</ul>
```{r}
plt3 <- ggplot() + geom_line(data=table[,c("Starttime","X.Packets","cluster")], aes(x=Starttime, y=X.Packets,color=cluster)) 
plt3 <- plt3 + ggtitle("Network Cluster Analysis", subtitle="K-means Security Results | K=2")
plt3 <- plt3 + xlab("Number of Hours (Cumulative by 5)") + ylab("Number of Packets")
plt3 <- plt3 + scale_colour_gradient(limits=c(1, 2), low="red", high="green")
plt3
```

<h2>Apply K-means on the data</h2>
<ul><li>At K=2</li></ul>
```{r}
set.seed(54321)
cluster <- kmeans(table[,2:ncol(table)], centers=10, nstart=30)
table$cluster <- cluster$cluster
```

<h2>K-means Results</h2>
<ul>
<li>At K=10</li>
<li>The initial security vulnerabilites were captured</li>
<li>The dead spot vulnerabilities were undermined as not as bad as they should be considered</li>
<li>The latter security vulnerabilities were for the most part also captured consistently</li>
</ul>
```{r}
plt4 <- ggplot() + geom_line(data=table[,c("Starttime","X.Packets","cluster")], aes(x=Starttime, y=X.Packets,color=cluster)) 
plt4 <- plt4 + ggtitle("Network Cluster Analysis", subtitle="K-means Security Results | K=10")
plt4 <- plt4 + xlab("Number of Hours (Cumulative by 5)") + ylab("Number of Packets")
plt4 <- plt4 + scale_colour_gradient(limits=c(1, 10), low="red", high="green")
plt4
```

<h2>How wrong was the K-means algorithm at predicting network security?</h2>

<ul>
<li>At K=2 --> 3.171672</li>
<li>At K=10 --> 4.392572</li>
</ul>
```{r}

MSE <- sum( (table$X.Security - table$cluster)^2 )/length(table$X.Security)
RMSE <- sqrt(MSE)
RMSE
```

<h2>Concluding Analysis</h2>
<ul>
<li>The rankings in the data gave 10 different heats for network security</li>
<li>The clustering algorithm did better predicting <em>Secure</em> and <em>Not Secure</em> rather than the 10 different levels of security</li>
</ul>
<br><br><br><br>
